---
title: "analyses_script"
output: pdf_document
date: "2023-08-03"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(patchwork)
library(GGally)
library(ggpubr)
library(ggBubbles)

```

Water parameters

```{r}
#load data spreadsheet into dat_water
dat_water <- read_excel("Data/YSA_waterData.xlsx", sheet = 1, 
                        range = cell_cols("A:K"))

names(dat_water) = c("location_num", "water_location","water_tempC","water_pH","water_ppM","water_uscm","Amonia_mgL","Turbidity_NTU","Nitrate_mgL","Natural","Flowing")

#remove celsius characters and conver to numeric
dat_water$water_tempC <- as.numeric(gsub(pattern = "'C", replacement = "", dat_water$water_tempC))

#replace missing vsalues with the lowest measurable value -0.05
dat_water$Nitrate_mgL[is.na(dat_water$Nitrate_mgL)] <- 0.25

#Plot for temp 'C
p1 <- dat_water %>% ggplot(aes(x = location_num, y = water_tempC)) + 
  geom_col(fill="#ffc87a") +
  labs(x = "site location", y = "'C", title="Water temperature") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))



#Plot for water pH
p2 <- dat_water %>% ggplot(aes(x = location_num, y = water_pH)) + 
  geom_col(fill="#86eee7") +
  labs(x = "site location", y = "pH",title="Water pH") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#Plot for water ppM
p3 <- dat_water %>% ggplot(aes(x = location_num, y = water_ppM)) + 
  geom_col(fill="#a3ecff") +
  labs(x = "site location", y = "ppM",title="Water mineral concentration") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#Plot for water μscm
p4 <- dat_water %>% ggplot(aes(x = location_num, y = water_uscm)) + 
  geom_col(fill="#a3d1ff") +
  labs(x = "site location", y = "us/cm",title="Water salt concentration") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#Plot for water Amonia_mgL
p5 <- dat_water %>% ggplot(aes(x = location_num, y = Amonia_mgL)) + 
  geom_col(fill="#13b469") +
  labs(x = "site location", y = "mg/L",title="Water ammonia") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#Plot for water Turbidity_NTU
p6 <- dat_water %>% ggplot(aes(x = location_num, y = Turbidity_NTU)) + 
  geom_col(fill="#52c6ff") +
  labs(x = "site location", y = "NTU",title="Water turbidity") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#Plot for water Nitrate_mgL
p7 <- dat_water %>% ggplot(aes(x = location_num, y = Nitrate_mgL)) + 
  geom_col(fill="#0091ff") +
  labs(x = "site location", y = "mg/L",title="Water nitrate") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#combine graphs into a figure
p1 + p2 + p6 +
  plot_layout(nrow = 3) + 
  plot_annotation(tag_levels = "A")
ggsave("water_plot1.png", width = 16, height = 12, units = "cm")

p3 + p4 + p5 + p7 +
  plot_layout(nrow = 4) + 
  plot_annotation(tag_levels = "A")

ggsave("water_plot2.png", width = 16, height = 16, units = "cm")

```

categorical graph
```{r}
#add columns
dat_water$Natural_Flowing <- paste(dat_water$Natural, dat_water$Flowing, sep = "_")
#make graph for categories
dat_water$Natural_Flowing_col <- case_when(dat_water$Natural_Flowing == "Natural_Flowing" ~ "#1B9E77",
                                            dat_water$Natural_Flowing == "Natural_Still" ~ "#66A61E",
                                            dat_water$Natural_Flowing == "Artificial_Flowing" ~ "#D95F02",
                                            dat_water$Natural_Flowing == "Artificial_Still" ~ "#CC6666")
  
p8 <- dat_water %>% ggplot(aes(x = Natural, y = Flowing, colour = location_num)) + 
  geom_point(position = position_surround(), size = 5) +
  labs(x = "Natural", y = "Flowing",title="Water types") +
  theme_classic() +
  scale_colour_manual(values = c("#1B9E77", "#1B9E77", "#1B9E77", "#1B9E77", "#1B9E77", "#CC6666", "#CC6666", "#D95F02", "#D95F02","#CC6666", "#D95F02", "#1B9E77", "#1B9E77", "#D95F02", "#CC6666", "#1B9E77", "#CC6666", "#D95F02", "#66A61E", "#66A61E")) + 
  theme(plot.title = element_text(hjust = 0.5), legend.key.height = unit(0.1, 'cm'))
ggsave("Natural_FlowingPlot.png", width = 16, height = 12, units = "cm")
```

pairwaise relationships among RG_ratio and water parameters
```{r}
ggpairs(dat_water, columns = 3:9) 

ggsave("water_corr.png", width = 12, height = 12, units = "cm", scale = 1.2)
```

Experiment 1

```{r}
#load data spreadsheet into dat_Original
dat_Original <- read_excel("Data/YSA_waterData.xlsx", sheet = 2, col_types=c("numeric","numeric","text"), range=cell_cols("A:C"))
names(dat_Original) = c("tea_num:", "weight","tea_type")
#dat_Original


#Finding initial mean and sd of R and G teabags in dat_Original
dat_Original %>% 
  group_by(tea_type) %>%
  summarise_at(vars(weight), list(mean = mean, sd = sd)) -> dat_Mean


#graph for original weight of teabags
means1 <- ggplot(dat_Mean, aes(x = tea_type, y = mean)) + 
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd), width=.3) +
  theme_classic() +
  geom_point(size = 2) + 
  ylim(0, 2.5) +
  labs(x = "Type of tea", y = "Weight [g]",title="Tea bag weight before") + 
  theme(plot.title = element_text(hjust = 0.5))

#combining graphs into a figure
means1 + means2
ggsave("MeansGraph.png", width = 16, height = 8, units = "cm")
```


```{r}
#load data spreadsheet into sheet_names
sheet_names <- excel_sheets("Data/TeaExperiment1_.xlsx")
sheet_names <- sheet_names[-1]
list_All <- lapply(sheet_names, function(x) {as.data.frame(read_excel("Data/TeaExperiment1_.xlsx", sheet = x))}) #read all sheets to list
names(list_All) <- sheet_names
dat_After <- do.call(rbind, list_All)
names(dat_After)
str(dat_After)

names(dat_After) <- c("weight2","tea_type2","info","location")
#transforming data into usable state
dat_After %>% 
  group_by(tea_type2) %>%
  summarise_at(vars(weight2), list(mean = mean, sd = sd)) %>%
  filter(complete.cases(.)) -> dat_Mean_A

dat_After %>% 
  group_by(tea_type2, .drop = FALSE) %>% count()

#graph for original weight of teabags
means2 <- ggplot(dat_Mean_A, aes(x = tea_type2, y = mean)) + 
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd), width=.3) +
  theme_classic() +
  geom_point(size = 2) + 
  ylim(0, 2.5) +
  labs(x = "Type of tea", y = "Weight [g]",title="Tea bag weight after") + 
  theme(plot.title = element_text(hjust = 0.5))



#percent weight change for G
100 - 0.6233881	/ 2.141744 * 100 #71% loss
#percent weight change for R
100 - 1.2017013 / 1.845925 * 100 #35% loss

#error propagation for G
0.6233881	/ 2.141744 * sqrt((0.04704403 /2.141744)^2 + (0.2581518	/0.6233881)^2 )*100 # 12.07029
#error propagation for R
1.2017013 / 1.845925 * sqrt((0.04811922	/1.845925)^2 + (0.2953956	/1.2017013 )^2 )*100 # 16.09231

#lnCVR effect size - change of variation on log scale:
#G
log((0.2581518 / 0.6233881) / (0.04704403/2.141744)) #2.93667
#R
log((0.2953956	/1.2017013) / (0.04811922	/1.845925)) #2.243876

```


#Site-specific effects
``` {r}
#load data spreadsheet into sheet_names
sheet_names <- excel_sheets("Data/TeaExperiment1_.xlsx")
sheet_names <- sheet_names[-1]
list_All <- lapply(sheet_names, function(x) {as.data.frame(read_excel("Data/TeaExperiment1_.xlsx", sheet = x))}) #read all sheets to list
names(list_All) <- sheet_names
dat_After <- do.call(rbind, list_All)
names(dat_After)
str(dat_After)
dim(dat_After)

names(dat_After) <- c("weight2","tea_type2","info","location")

dat_After %>% 
  group_by(location, .drop = FALSE) %>%
  count() #just counts all rows

#count bags with holes
dat_After %>% filter(info == "holes") %>% count() #63
dat_After$info

#remove all bags with holes or no weights
dat_After2 <- dat_After %>% 
  filter(is.na(info) | info != "holes") %>%
  filter(!is.na(weight2))
  
dim(dat_After2) #144
dat_After2$weight2
  
#count by site
dat_After2 %>% 
  group_by(location) %>%
  count()

144/16

table(dat_After2$tea_type2, dat_After2$location) 

names(dat_After2)  
#calculating mean and sd of data
dat_After2 %>% 
  group_by(location, tea_type2) %>%
  summarise_at(vars(weight2), list(mean = mean, sd = sd))

print(dat_After2)

dat_After2 %>% 
  drop_na(weight2) %>%
  ggplot(aes(x = location,
             y = weight2,
             fill = tea_type2)) +
  #geom_violin() +
  geom_boxplot(varwidth = TRUE) +
  theme_classic() +
  scale_fill_manual(values = c("springgreen3", "#FF6070")) +
  labs(x = "site location", y = "Tea bag weight [g]") + 
  geom_vline(xintercept = seq(0.5, 16, by = 1), color="gray", size=.5, alpha=.5) + # set vertical lines between x groups
  labs(title="Tea bag weight after 90 days") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("Exp1WeightGraph.png", width = 12, height = 8, units = "cm", scale = 1.2)

help(geom_boxplot)

#find sites to be removed
dat_After2 %>% 
  group_by(location, tea_type2) %>%
  count() %>% 
  filter(n < 3) %>% 
  select(location) -> remove_sites

remove_sites <- unique(remove_sites$location) #vector of codes

#reduce data frame
dat_After3 <- subset(dat_After2, !(location %in% remove_sites))
dim(dat_After3) #105

#change to factor and column name before joining
dat_Mean$tea_type2 <- as.factor(dat_Mean$tea_type)

# add column with initial average weights
dat_After3 <- left_join(dat_After3, dat_Mean, by = "tea_type2")

#calculate percent mass loss separately for R and G
dat_After3$pct_change <- (dat_After3$weight2 - dat_After3$mean) / dat_After3$mean * 100

names(dat_After3)
#plot pct_change
dat_After3 %>%
  ggplot(aes(x = location,
             y = pct_change,
             fill = tea_type)) +
  geom_boxplot(varwidth = TRUE) +
  theme_classic() +
  scale_fill_manual(values = c("springgreen3", "#FF6070")) +
  labs(x = "site location", y = "Tea bag weight change %") + 
  geom_vline(xintercept = seq(0.5, 16, by = 1), color="gray", size=.5, alpha=.5) + # set vertical lines between x groups
  labs(title="Tea bag weight % change after 90 days") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("Exp1SiteWeight.png", width = 12, height = 8, units = "cm", scale = 1.2)

#calculate ratio of G to R
names(dat_After3)

dat_After3 %>% 
  group_by(location, tea_type) %>%
  summarise(pctmean = mean(pct_change)) -> dat_After3_means
dat_After3_means

dat_After3_means %>% group_by(location) %>% pivot_wider(names_from = tea_type, values_from = pctmean) -> dat_After3_means_wide

dat_After3_means_wide$RG_ratio <- dat_After3_means_wide$R / dat_After3_means_wide$G

#barplot of RG_ratio


dat_After3_means_wide %>% ggplot(aes(x = location, y = RG_ratio)) + 
  geom_col() +
  labs(x = "site location", y = "% loss ratio of Rooibos to Green tea",title="Ratio of Percentage Change") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("figure8.png", width = 12, height = 6, units = "cm", scale = 1.2)

```

Join decomposotion and water data tables
```{r}
dat_water <- dat_water %>% rename("location" = "location_num")
dat_After4 <- left_join(ungroup(dat_After3_means_wide), dat_water)
#str(dat_After4)
```

G percent decomposition and water characteristics

```{r}
#making graphs for different water characteristics
g1 <- dat_After4 %>% ggplot(aes(x = water_tempC, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "% Change", title = "Effect of temp on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g2 <- dat_After4 %>% ggplot(aes(x = water_pH, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "% Change", title = "Effect of pH on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g3 <- dat_After4 %>% ggplot(aes(x = Turbidity_NTU, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "% Change", title = "Effect of turbidity on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g4 <- dat_After4 %>% ggplot(aes(x = water_ppM, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "% Change", title = "Effect of minerals on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g5 <- dat_After4 %>% ggplot(aes(x = Amonia_mgL, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "% Change", title = "Effect of amonia on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g6 <- dat_After4 %>% ggplot(aes(x = Nitrate_mgL, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "% Change", title = "Effect of nitrate on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#Combining graphs into a figure
g1 + g2 + g3 + g4 + g5 + g6 +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")

ggsave("water_G_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```

G percent decomposition and water characteristics

```{r}
str(dat_After4)
#making graphs of correlation between water characteristics and decomposition 
r1 <- dat_After4 %>% ggplot(aes(x = water_tempC, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "% Change", title = "Effect of temp on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r2 <- dat_After4 %>% ggplot(aes(x = water_pH, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "% Change", title = "Effect of pH on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r3 <- dat_After4 %>% ggplot(aes(x = Turbidity_NTU, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "% Change", title = "Effect of turbidity on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r4 <- dat_After4 %>% ggplot(aes(x = water_ppM, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "% Change", title = "Effect of minerals on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r5 <- dat_After4 %>% ggplot(aes(x = Amonia_mgL, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "% Change", title = "Effect of amonia on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r6 <- dat_After4 %>% ggplot(aes(x = Nitrate_mgL, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "% Change", title = "Effect of nitrate on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#combining graphs into a figure
r1 + r2 + r3 + r4 + r5 + r6 +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")
ggsave("water_R_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```

RG ratio decomposition index and water characteristics

Join decomposotion and water data tables
```{r}
#Graphing correlation between RG ratio and decomposition
w1 <- dat_After4 %>% ggplot(aes(x = water_tempC, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "RG ratio", title = "Effect of temp on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w2 <- dat_After4 %>% ggplot(aes(x = water_pH, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "RG ratio", title = "Effect of pH on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w3 <- dat_After4 %>% ggplot(aes(x = Turbidity_NTU, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "RG ratio", title = "Effect of turbidity on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w4 <- dat_After4 %>% ggplot(aes(x = water_ppM, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "RG ratio", title = "Effect of minerals on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w5 <- dat_After4 %>% ggplot(aes(x = Amonia_mgL, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "RG ratio", title = "Effect of amonia on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w6 <- dat_After4 %>% ggplot(aes(x = Nitrate_mgL, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "RG ratio", title = "Effect of nitrate on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#combining graphs into a figure
w1 + w2 + w3 + w4 + w5 + w6 +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")
ggsave("water_RG_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```


Experiment 2

```{r}
#getting data from spreadsheet into experiment2
experiment2 <- read_excel("Data/EXPERIMENT2.xlsx")
names(experiment2) = c("Bag_number", "Tea_Type", "Weight_Before", "Weight_After","site2")


#Finding mean and sd of R and G teabags in experiment2
experiment2 %>% 
  group_by(Tea_Type) %>%
  summarise_at(vars(Weight_Before), list(mean = mean, sd = sd)) -> exp_Mean
experiment2 %>% 
  group_by(Tea_Type) %>%
  summarise_at(vars(Weight_After), list(mean = mean, sd = sd)) -> exp_Mean2


#graph for original weight of teabags
mean1 <- ggplot(exp_Mean, aes(x = Tea_Type, y = mean)) + 
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd), width=.3) +
  theme_classic() +
  geom_point(size = 2) + 
  ylim(0, 2.5) +
  labs(x = "Type of tea", y = "Weight [g]",title="Tea bag weight before") + 
  theme(plot.title = element_text(hjust = 0.5))

#graph for after weight of teabags
mean2 <- ggplot(exp_Mean2, aes(x = Tea_Type, y = mean)) + 
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd), width=.3) +
  theme_classic() +
  geom_point(size = 2) + 
  ylim(0, 2.5) +
  labs(x = "Type of tea", y = "Weight [g]",title="Tea bag weight after") + 
  theme(plot.title = element_text(hjust = 0.5))
#combining graphs into figure
mean1 + mean2
ggsave("MeansGraph2.png", width = 16, height = 8, units = "cm")
```

```{r}

#percent weight change for G
100 - 0.6233881	/ 2.141744 * 100 #71% loss
#percent weight change for R
100 - 1.2017013 / 1.845925 * 100 #35% loss

#error propagation for G
0.6233881	/ 2.141744 * sqrt((0.04704403 /2.141744)^2 + (0.2581518	/0.6233881)^2 )*100 # 12.07029
#error propagation for R
1.2017013 / 1.845925 * sqrt((0.04811922	/1.845925)^2 + (0.2953956	/1.2017013 )^2 )*100 # 16.09231

#lnCVR effect size - change of variation on log scale:
#G
log((0.2581518 / 0.6233881) / (0.04704403/2.141744)) #2.93667
#R
log((0.2953956	/1.2017013) / (0.04811922	/1.845925)) #2.243876

```

#Site-specific effects
``` {r}
experiment2 %>% 
  group_by(site2, .drop = FALSE) %>%
  count() #just counts all rows
  
dim(experiment2) #400
experiment2$Weight_Before
experiment2$Weight_After

table(experiment2$Tea_Type, experiment2$site2) 

names(experiment2)  
#getting mean and sd of experiment2 samples before and after experiment
experiment2 %>% 
  group_by(site2, Tea_Type) %>%
  summarise_at(vars(Weight_Before), list(mean = mean, sd = sd))

experiment2 %>% 
  group_by(site2, Tea_Type) %>%
  summarise_at(vars(Weight_After), list(mean = mean, sd = sd))

#making boxplot for gram change of samples from decomposition
experiment2 %>% 
  drop_na(Weight_After) %>%
  ggplot(aes(x = site2,
             y = Weight_After,
             fill = Tea_Type)) +
  geom_boxplot(varwidth = TRUE) +
  theme_classic() +
  scale_fill_manual(values = c("springgreen3", "#FF6070")) +
  labs(x = "site location", y = "Tea bag weight [g]") + 
  geom_vline(xintercept = seq(0.5, 21, by = 1), color="gray", size=.5, alpha=.5) + # set vertical lines between x groups
  labs(title="Tea bag weight after 90 days") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("Exp2WeightGraph.png", width = 12, height = 8, units = "cm", scale = 1.2)

#calculate percent mass loss separately for R and G
experiment2$pct_change <- (experiment2$Weight_After - experiment2$Weight_Before) / experiment2$Weight_Before * 100

#check outliers
experiment2 %>% 
  filter(pct_change > -5) %>% View()


#plot for pct_change of samples 
experiment2 %>%
  ggplot(aes(x = site2,
             y = pct_change,
             fill = Tea_Type)) +
  geom_boxplot(varwidth = TRUE) +
  theme_classic() +
  scale_fill_manual(values = c("springgreen3", "#FF6070")) +
  labs(x = "site location", y = "Tea bag weight change %") + 
  geom_vline(xintercept = seq(0.5, 21, by = 1), color="gray", size=.5, alpha=.5) + # set vertical lines between x groups
  labs(title="Tea bag weight % change after 90 days") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("Exp2SiteWeight.png", width = 12, height = 8, units = "cm", scale = 1.2)



#calculate ratio of G to R - using mean values of R and G percent change
names(experiment2)
names(dat_After3)

experiment2 %>% 
  group_by(site2, Tea_Type) %>%
  summarise(pctmean = mean(pct_change)) -> experiment2_means
dexperiment2_means

experiment2_means %>% group_by(site2) %>% pivot_wider(names_from = Tea_Type, values_from = pctmean) -> experiment2_means_wide

experiment2_means_wide$RG_ratio <- experiment2_means_wide$R / experiment2_means_wide$G

#barplot of RG_ratio

experiment2_means_wide %>% ggplot(aes(x = site2, y = RG_ratio)) + 
  geom_col() +
  labs(x = "site location", y = "% loss ratio of Rooibos to Green tea",title="Ratio of Percentage Change") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("figureExp2.png", width = 12, height = 6, units = "cm", scale = 1.2)

```

Join decomposotion and water data tables
```{r}
names(experiment2_means_wide)
names(dat_water)
dat_water$location <- dat_water$location_num
experiment2_means_wide <- experiment2_means_wide %>% rename("location" = "site2")
experiment2_means_wide <- left_join(ungroup(experiment2_means_wide), dat_water)
names(experiment2_means_wide)
```

G percent decomposition and water characteristics

```{r}
#making plots for correlation between green decomposition and water parameters
g21 <- experiment2_means_wide %>% ggplot(aes(x = water_tempC, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "% Change", title = "Effect of temp on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g22 <- experiment2_means_wide %>% ggplot(aes(x = water_pH, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "% Change", title = "Effect of pH on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g23 <- experiment2_means_wide %>% ggplot(aes(x = Turbidity_NTU, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "% Change", title = "Effect of turbidity on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g24 <- experiment2_means_wide %>% ggplot(aes(x = water_ppM, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "% Change", title = "Effect of minerals on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g25 <- experiment2_means_wide %>% ggplot(aes(x = Amonia_mgL, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "% Change", title = "Effect of amonia on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

g26 <- experiment2_means_wide %>% ggplot(aes(x = Nitrate_mgL, y = -G)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "% Change", title = "Effect of nitrate on green tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#combining graphs together to create figure
g21 + g22 + g23 + g24 + g25 + g26 +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")

ggsave("Exp2G_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```

R percent decomposition and water characteristics

```{r}
#graph for rooibos weight correlation to water parameters
r21 <- experiment2_means_wide %>% ggplot(aes(x = water_tempC, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "% Change", title = "Effect of temp on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r22 <- experiment2_means_wide %>% ggplot(aes(x = water_pH, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "% Change", title = "Effect of pH on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r23 <- experiment2_means_wide %>% ggplot(aes(x = Turbidity_NTU, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "% Change", title = "Effect of turbidity on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r24 <- experiment2_means_wide %>% ggplot(aes(x = water_ppM, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "% Change", title = "Effect of minerals on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r25 <- experiment2_means_wide %>% ggplot(aes(x = Amonia_mgL, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "% Change", title = "Effect of amonia on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

r26 <- experiment2_means_wide %>% ggplot(aes(x = Nitrate_mgL, y = -R)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "% Change", title = "Effect of nitrate on rooibos tea decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#combining graphs into figure
r21 + r22 + r23 + r24 + r25 + r26 +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")
ggsave("Exp2R_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```

RG ratio decomposition index and water characteristics

```{r}
#making plots for correlation between RG ratio and water parameters
w21 <- experiment2_means_wide %>% ggplot(aes(x = water_tempC, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "RG ratio", title = "Effect of temp on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w22 <- experiment2_means_wide %>% ggplot(aes(x = water_pH, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "RG ratio", title = "Effect of pH on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w23 <- experiment2_means_wide %>% ggplot(aes(x = Turbidity_NTU, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "RG ratio", title = "Effect of turbidity on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w24 <- experiment2_means_wide %>% ggplot(aes(x = water_ppM, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "RG ratio", title = "Effect of minerals on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w25 <- experiment2_means_wide %>% ggplot(aes(x = Amonia_mgL, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "RG ratio", title = "Effect of amonia on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w26 <- experiment2_means_wide %>% ggplot(aes(x = Nitrate_mgL, y = RG_ratio)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "RG ratio", title = "Effect of nitrate on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#combining graphs into figure
w21 + w22 + w23 + w24 + w25 + w26 +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")
ggsave("Exp2RG_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```

### Compare means from Experiment1 and Experimen2

```{r}
#merge by location

experiments12 <- left_join(dat_After3_means_wide, experiment2_means_wide, by = c("location"))
names(experiments12)
#graph for comparison between experiment1 and experiment2 mean % change in green tea
exp12G <- experiments12 %>% 
  ggplot(aes(x = -G.x, y = -G.y)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "experiment1 mean % change", y = "experiment2 mean % change", title = "Experiment1 vs Experiment2 green tea decomposition") +
  geom_smooth(method = lm) +
  #ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

experiments12 <- left_join(dat_After3_means_wide, experiment2_means_wide, by = c("location"))
names(experiments12)
#graph for comparison between experiment1 and experiment2 mean % change in rooibos tea
exp12R <- experiments12 %>% 
  ggplot(aes(x = -R.x, y = -R.y)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "experiment1 mean % change", y = "experiment2 mean % change", title = "Experiment1 vs Experiment2 rooibos tea decomposition") +
  geom_smooth(method = lm) +
  #ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#graph for comparison between experiment1 and experiment2 mean % change in RG ratio
exp12RG <- experiments12 %>% 
  ggplot(aes(x = -RG_ratio.x, y = -RG_ratio.y)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "experiment1 mean ratio % changes", y = "experiment2 mean ratio % changes", title = "Experiment1 vs Experiment2 rooibos to green ratio") +
  geom_smooth(method = lm) +
  #ylim(0, 100) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)
#combining graphs toghether into figure
exp12RG + exp12G + exp12R +
  plot_layout(nrow = 3, ncol = 1) + 
  plot_annotation(tag_levels = "A")
ggsave("Exp1to2cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
  
  
  
```












##################### Explore catgorical water types

RG ratio decomposition index and water characteristics

By natural or artificial sites for water
```{r}
w21a <- experiment2_means_wide %>% ggplot(aes(x = water_tempC, y = RG_ratio, color = Natural)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "RG ratio", title = "Effect of temp on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w22a <- experiment2_means_wide %>% ggplot(aes(x = water_pH, y = RG_ratio, color = Natural)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "RG ratio", title = "Effect of pH on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w23a <- experiment2_means_wide %>% ggplot(aes(x = Turbidity_NTU, y = RG_ratio, color = Natural)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "RG ratio", title = "Effect of turbidity on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w24a <- experiment2_means_wide %>% ggplot(aes(x = water_ppM, y = RG_ratio, color = Natural)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "RG ratio", title = "Effect of minerals on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w25a <- experiment2_means_wide %>% ggplot(aes(x = Amonia_mgL, y = RG_ratio, color = Natural)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "RG ratio", title = "Effect of amonia on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w26a <- experiment2_means_wide %>% ggplot(aes(x = Nitrate_mgL, y = RG_ratio, color = Natural)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "RG ratio", title = "Effect of nitrate on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w21a + w22a + w23a + w24a + w25a + w26a +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")
ggsave("Exp2RGa_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```

By flowing versus still water
```{r}
w21b <- experiment2_means_wide %>% ggplot(aes(x = water_tempC, y = RG_ratio, color = Flowing)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water Temp", y = "RG ratio", title = "Effect of temp on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w22b <- experiment2_means_wide %>% ggplot(aes(x = water_pH, y = RG_ratio, color = Flowing)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water pH", y = "RG ratio", title = "Effect of pH on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w23b <- experiment2_means_wide %>% ggplot(aes(x = Turbidity_NTU, y = RG_ratio, color = Flowing)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water NTU", y = "RG ratio", title = "Effect of turbidity on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w24b <- experiment2_means_wide %>% ggplot(aes(x = water_ppM, y = RG_ratio, color = Flowing)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water ppM", y = "RG ratio", title = "Effect of minerals on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w25b <- experiment2_means_wide %>% ggplot(aes(x = Amonia_mgL, y = RG_ratio, color = Flowing)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water amonia mg/L", y = "RG ratio", title = "Effect of amonia on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w26b <- experiment2_means_wide %>% ggplot(aes(x = Nitrate_mgL, y = RG_ratio, color = Flowing)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "Water nitrate mg/L", y = "RG ratio", title = "Effect of nitrate on relative decomposition") +
  geom_smooth(method = lm) +
  ylim(0, 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)

w21b + w22b + w23b + w24b + w25b + w26b +
  plot_layout(nrow = 3, ncol = 2) + 
  plot_annotation(tag_levels = "A")
ggsave("Exp2RGb_cor.png", width = 16, height = 12, units = "cm", scale = 1.6)
```